---
title: "Exploratory Data Analysis Project 2"
author: "Halley Wang"
date: "September 4, 2015"
output: html_document
---



```{r golbal_options, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, result = "asis", message=FALSE, fig.height=7, fig.width=12)
```

```{r load_package, results = "hide"}
## Loading the necessary packages
library(ggplot2)

```

```{r load data, cache = TRUE}
SCC <- readRDS("Source_Classification_Code.rds")
NEI <- readRDS("summarySCC_PM25.rds")
```

1. Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Using the base plotting system, make a plot showing the total PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.

```{r question1, cache = TRUE}
# tapply method. however this method can hardly retrieve the year element. Have to input it manually.
emissionSum = with(NEI, tapply(Emissions, year, sum))
yearunique = unique(NEI$year)
emissionT1 = data.frame(year = yearunique, emission = emissionSum)

# Initial the points
with(emissionT1, plot(year, emission))
# Connect by line:
with(emissionT1, lines(year, emission))
# Add a title:
title("Total PM2.5 Emmissions in US(1999~2008)")
```

2. Have total emissions from PM2.5 decreased in the Baltimore City, Maryland (fips == 24510) from 1999 to 2008? Use the base plotting system to make a plot answering this question.

```{r question2, cache = TRUE}
# aggregate method. Easy and feasible.
BaltimoreT = subset(NEI, fips == 24510)
emissionT2 = aggregate(Emissions~year, data = BaltimoreT, sum)

# Initial the points
with(emissionT2, plot(year, Emissions))
# Connect by line:
with(emissionT2, lines(year, Emissions))
# Add a title:
title("Total PM2.5 Emmissions in Baltimore(1999~2008)")

```


3. Of the four types of sources indicated by the type (point, nonpoint, onroad, nonroad) variable, which of these four sources have seen decreases in emissions from 1999-2008 for Baltimore City? Which have seen increases in emissions from 1999-2008? Use the ggplot2 plotting system to make a plot answer this question.

```{r question3, cache = TRUE}

emissionT3 = aggregate(Emissions~year+type, data = BaltimoreT, sum)
# Establish the ggplot object
p = ggplot(emissionT3, aes(year, Emissions))

# Add point and line layers(Separating type by colors)
# Add facets along the types
# Add title
p + geom_point()+geom_line(aes(color = type)) + facet_grid( .~type) + labs(title = "Emissions under different types from 1999~2008 in Baltimore")
```





4. Across the United States, how have emissions from coal combustion-related sources changed from 1999-2008?







5. How have emissions from motor vehicle sources changed from 1999-2008 in Baltimore City?









6. Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in Los Angeles County, California (fips == 06037). Which city has seen greater changes over time in motor vehicle emissions?








